---
title: "SQL"
output: html_document
---

## Filtering Columns and Rows

With SQL, you can filter columns and rows by using SELECT and WHERE. The following is an example using the Lahman database. 
First, however, we need to load both the Lahman database and package sqldf.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

###Example One:
What were the homerun totals for the 1927 Yankees? 
I could write the following query: 

```{r}
query<-"select playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA'and yearid=1927"
sqldf(query)
```
###Example Two:
Find all instances where the Yankees have hit 40 homeruns or more.
```{r}

query<-"SELECT playerID, teamID, HR FROM Batting
WHERE teamID='NYA' and HR>=40"
sqldf(query)
``` 

###Example Three:
Find all instances where a player had more than 40 homeruns but less than 60 strikeouts.
```{r}

query<-"SELECT playerID, teamID, HR, SO FROM Batting
WHERE HR>40 and SO<60"
sqldf(query)
```
###Example Four:
Find all instances of Phillies in the 1970's hitting more than 30 homeruns.
```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting
WHERE teamID='PHI' and  HR>30 and yearID>1969 and yearID<1980"
sqldf(query)
```




## ORDER BY

With SQL, you can order the rows of your output. 

###Example One: 
Find every instance of a player hitting more than 50 homerun, and put the players with the most homeruns at the top. 

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```
###Example Two:
Find all instances of a player striking out less than 10 times, with each player having at least 400 at-bats. The players with the least amount of strikeouts should be at the top.
```{r}
query<-"SELECT playerID, teamID, SO, AB FROM Batting
WHERE SO<10 and AB>=400 
ORDER BY SO"
sqldf(query)
```

