---
title: "SQL"
output: html_document
---

## Filtering Columns and Rows

With SQL, you can filter columns and rows by using SELECT and WHERE. The following is an example using the Lahman database. 
First, however, we need to load both the Lahman database and package sqldf.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

###Example One:
What were the homerun totals for the 1927 Yankees? 
I could write the following query: 

```{r}
query<-"select playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA'and yearid=1927"
sqldf(query)
```
###Example Two:
Find all instances where the Yankees have hit 40 homeruns or more.
```{r}

query<-"SELECT playerID, teamID, HR FROM Batting
WHERE teamID='NYA' and HR>=40"
sqldf(query)
``` 

###Example Three:
Find all instances where a player had more than 40 homeruns but less than 60 strikeouts.
```{r}

query<-"SELECT playerID, teamID, HR, SO FROM Batting
WHERE HR>40 and SO<60"
sqldf(query)
```
###Example Four:
Find all instances of Phillies in the 1970's hitting more than 30 homeruns.
```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting
WHERE teamID='PHI' and  HR>30 and yearID>1969 and yearID<1980"
sqldf(query)
```




## ORDER BY

With SQL, you can order the rows of your output. 

###Example One: 
Find every instance of a player hitting more than 50 homerun, and put the players with the most homeruns at the top. 

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```
###Example Two:
Find all instances of a player striking out less than 10 times, with each player having at least 400 at-bats. The players with the least amount of strikeouts should be at the top.
```{r}
query<-"SELECT playerID, teamID, SO, AB FROM Batting
WHERE SO<10 and AB>=400 
ORDER BY SO"
sqldf(query)
```




##AGGREGATION
Aggregation is grouping different information, essentially 'squashing' them together. This can be done using the GROUP BY command. 

###Example One:
Find the homerun totals for Babe Ruth. 

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting 
WHERE playerID='ruthba01'
GROUP BY playerID"
sqldf(query)
```
###Example Two:
Find career homerun totals for all players but list only those with a total of 600 or more, and put in desecending order. 

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"
sqldf(query)
```

###Example Three:
What player have the highest average per season homerun totals? List only those with averages greater than 30, and order from highest to lowest. 

```{r}
query<-"SELECT playerID, avg(HR) FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"
sqldf(query)
```


##Joining
By using the commands INNER JOIN and ON, tables can be added together to create larger data output options. 

###Example One
First and last name, along with team, year, and homeruns for Babe Ruth. 

```{r}
query<-"SELECT nameFIRST,nameLAST,teamID,yearID,HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"
sqldf(query)
```

###Example Two:
Find all instances of players getting more than 50 homeruns starting with the greatest number. Include first and last names, team, year, and home runs. 

```{r}
query<-"SELECT nameFIRST, nameLAST, teamID, yearID, HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```


###Example Three:
Find Babe Ruth's homeruns by season, with the team name displayed.

```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR
FROM Batting INNER JOIN Teams ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID WHERE playerID='ruthba01'"
sqldf(query)
```

###Example Four:
Babe Ruth: First and last name, team name, year, and homeruns.

```{r}
query<-"SELECT nameFirst,nameLast,name,Batting.yearID,Batting.HR 
FROM (Batting INNER JOIN Master
ON Batting.playerID=Master.playerID) INNER JOIN Teams
ON Batting.teamID=Teams.teamID AND Batting.yearID=Teams.yearID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```
###Example Five:
Find career homerun totals for players who have hit at least 600 homeruns. The highest homeruns should be at the top, while also including first name, last name, and homerun total.

```{r}
query<-"SELECT nameFirst,nameLast,sum(HR)
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
GROUP BY Batting.playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"

sqldf(query)
```
```{r}
query<-"SELECT nameFirst,nameLast,avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerID=Master.playerID
WHERE nameFirst='Bob'
GROUP BY Salaries.playerID
HAVING avg(salary)>1000000
ORDER BY avg(salary) DESC"

sqldf(query)
```

